Bringing Expectations to the Collective Bargaining Table: Evidence from Brazilian Firms

### 2021-04-22 Drafts/Draft_1.tex
Citations edit

\documentclass[12pt]{article}
\usepackage[capposition=top]{floatrow}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{float}
\usepackage{floatrow}
\usepackage{eurosym}
\usepackage{csvsimple}
\usepackage{booktabs}
\usepackage{tikz}
\usepackage{hyperref}
\usepackage{adjustbox}
\usepackage{ulem}
\usepackage{caption}
\usepackage{tablefootnote}
\usepackage{ntheorem}
\usepackage{sectsty}
\usepackage[justification=centering,textfont={sc},labelfont={rm}]{caption}
\usepackage{geometry}
\geometry{margin=1in}
\usepackage[numbers, square]{natbib}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage[section]{placeins}
\onehalfspacing
\begin{document}
	\title{Bringing Expectations to the Collective Bargaining Table: Evidence from Brazilian Firms}
	\author{Valerie R. Boctor\thanks{This project is a collaboration with Roberto Hsu Rocha. This document, including all tables and figures, was written by Valerie Boctor. All mistakes are my own. }}
	\date{\today}
	\maketitle

@@

			\item What are the implications of wage rigidities and employment behavior for optimal monetary policy?
		\end{enumerate}
% =================================================================================================
% 												BACKGROUND	
% =================================================================================================
	\section{Background}
	In this section, we discuss our intended contributions to the literature and provide key information about collective bargaining institutions in Brazil.
		\subsection{Literature Review}
		Olivei and Tenreyro (2007) posit that monetary policy shocks in the US have larger and less persistent effects during the first and second quarters than in the latter half of the year do to the timing of wage contract negotiations \cite{oliveiTimingMonetaryPolicy2007}. Because negotiations are more likely to occur in the first two quarters in the US, wages are more flexible during this period and respond more readily to monetary policy shocks. Since wages are less likely to be renegotiated during the third and fourth quarters, the response to monetary policy shocks during these periods tends to be smaller and more protracted. Our research builds on Olivei and Tenreyro's findings by measuring precisely the effect of an aggregate shock on firm-level wages, taking into account the relative timing of each firm's most recent CBA. If the reduction in nominal wages and/or employment is relatviely larger and less persistent for firms CBAs just after the news shock, then our results would corroborate Olivei and Tenreyo's aggregate findings.\footnote{See ``CBAs in Brazil'' for discussion of legally mandated downward nominal wage rigidity and workarounds.}

		\begin{itemize}
			\item Discuss David Card findings on unions and wage outcomes \cite{cardUnexpectedInflationReal1990}, \cite{cardEffectUnionsStructure1996}, \cite{cardStrikesBargainingSurvey1990}
			\item Discuss relevance to literature on nominal rigidities and macro models \cite{taylorStayingPowerStaggered2016} \cite{taylorChapter15Staggered1999} \cite{fukuiTheoryWageRigidity}
		\end{itemize}

		\subsection{Collective Bargaining Agreements in Brazil}
		In Brazil, collective bargaining agreements typically last 1 or 2 years. All CBA contracts have a maximum duration of 2 years, but can be extended for additional years. \cite{lagosLaborMarketInstitutions2021}. CBAs typically set a fixed nominal wage levels for the duration of the contract. It is important to note that nominal wages are legally required to be downward rigid. That is, nominal wages cannot be reduced unless the union approves the reduction first. Without an affirmation from the workers' union, if nominal wages fall, firms are legally required to offer workers additional hours. Conversely, a reduction in workers' hours must be compensated by a commensurate increase in the nominal wage rate. \cite{EnglishTranslationConsolidation2010}
		Given the constraints facing firms, wages (measured in average monthly remuneration) may still be reduced through separations, or by reclassifying workers to occupations with lower nominal wage levels.

@@

		Furthermore, we can rule out the possibility that the distribution of CBAs over time is driven by any particular industry. As Lagos (2021) points out, there is no bunching by certain industries around any particular month for CBAs \cite{lagosLaborMarketInstitutions2021}
% =================================================================================================
% 													DATA	
% =================================================================================================
	\section{Data}
		\subsection{RAIS}
		The Annual Social Information Report (RAIS) is an administrative dataset containing detailed information about workers and employers at the monthly frequency. The key variables of interest are monthly remuneration, employment status, 6-digit occupation codes (CBO), and 2-digit broad sector classifications from IBGE. We use average monthly remuneration by firm as the main wage variable throughout the study. We use the count of employed workers in a given month as the main employment variable of interest. 
		\begin{figure}
			\centering
			\caption{INSERT RAIS SUM STATS}
			\label{tab:SumStatsRAIS}
		\end{figure}

@@

% =================================================================================================
% 													THEORY	
% =================================================================================================
	\section{Future Research}
	Further analysis is required to pin down what factors are driving the reduction in nominal wages across all firms during the 2015-2016 recession. \ref{fig:NormWages} One possible explanation is that firms separate from workers during the recession, curtailing the workers' monthly earnings. There is some preliminary evidence of this in Figure \ref{fig:NormEmp}. Another potentially complementary explanation is that firms reclassify workers to occupations with lower implied nominal wages. While we have not yet carried out the analyses to corroborate whether the latter explanation is pertinent, we may have sufficient data to accomplish this in the near future.

	Another salient question is why firms that negotiate wages just after the shock set higher nominal wages than firms negotiating just before. \ref{fig:RegressionResults} One possible explanation is that firms in the post-shock group have higher levels of optimism about the future state of the economy and their own future profit margins, which may lead them to set higher wage levels. We may gain access to microdata on firms' own-price and own-cost expectations that would allow us to study the link between individual firms' expectations and negotiated wages as well as hiring and firing decisions.

	Following Olivei and Tenreyo (2007), we may adapt a dynamic stochastic general equilibrium model with uneven staggering of wage contracts and downward rigidity. \cite{oliveiTimingMonetaryPolicy2007}

	Another approach would be to adapt a DMP model to include inflation expectations, similar in spirit to Fukui (2020).  \cite{fukuiTheoryWageRigidity}

>>>>

\documentclass[12pt]{article}
\usepackage[capposition=top]{floatrow}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{float}
\usepackage{floatrow}
\usepackage{eurosym}
\usepackage{csvsimple}
\usepackage{booktabs}
\usepackage{tikz}
\usepackage{hyperref}
\usepackage{adjustbox}
\usepackage{ulem}
\usepackage{caption}
\usepackage{tablefootnote}
\usepackage{ntheorem}
\usepackage{sectsty}
\usepackage[justification=centering,textfont={sc},labelfont={rm}]{caption}
\usepackage{geometry}
\geometry{margin=1in}
\usepackage{natbib}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage[section]{placeins}
\onehalfspacing
\begin{document}
	\title{Bringing Expectations to the Collective Bargaining Table: Evidence from Brazilian Firms}
	\author{Valerie R. Boctor\thanks{This project is a collaboration with Roberto Hsu Rocha. This document, including all tables and figures, was written by Valerie Boctor. All mistakes are my own. }}
	\date{\today}
	\maketitle

@@

			\item What are the implications of wage rigidities and employment behavior for optimal monetary policy?
		\end{enumerate}
% =================================================================================================
% 												BACKGROUND	
% =================================================================================================
	\section{Background}
	In this section, we discuss our intended contributions to the literature and provide key information about collective bargaining institutions in Brazil.
		\subsection{Literature Review}
		Olivei and Tenreyro (2007) posit that monetary policy shocks in the US have larger and less persistent effects during the first and second quarters than in the latter half of the year do to the timing of wage contract negotiations. Because negotiations are more likely to occur in the first two quarters in the US, wages are more flexible during this period and respond more readily to monetary policy shocks. Since wages are less likely to be renegotiated during the third and fourth quarters, the response to monetary policy shocks during these periods tends to be smaller and more protracted. Our research builds on Olivei and Tenreyro's findings by measuring precisely the effect of an aggregate shock on firm-level wages, taking into account the relative timing of each firm's most recent CBA. If the reduction in nominal wages and/or employment is relatviely larger and less persistent for firms CBAs just after the news shock, then our results would corroborate Olivei and Tenreyo's aggregate findings.\footnote{See ``CBAs in Brazil'' for discussion of legally mandated downward nominal wage rigidity and workarounds.}

		% \begin{itemize}
		% 	\item Discuss David Card findings on unions and wage outcomes \cite{cardUnexpectedInflationReal1990}, \cite{cardEffectUnionsStructure1996}, \cite{cardStrikesBargainingSurvey1990}
		% 	\item Discuss relevance to literature on nominal rigidities and macro models \cite{taylorStayingPowerStaggered2016} \cite{taylorChapter15Staggered1999} \cite{fukuiTheoryWageRigidity}
		% \end{itemize}

		\subsection{Collective Bargaining Agreements in Brazil}
		In Brazil, collective bargaining agreements typically last 1 or 2 years. All CBA contracts have a maximum duration of 2 years, but can be extended for additional years. \cite{lagosLaborMarketInstitutions2021} CBAs typically set a fixed nominal wage levels for the duration of the contract. It is important to note that nominal wages are legally required to be downward rigid. That is, nominal wages cannot be reduced unless the union approves the reduction first. Without an affirmation from the workers' union, if nominal wages fall, firms are legally required to offer workers additional hours. Conversely, a reduction in workers' hours must be compensated by a commensurate increase in the nominal wage rate. \cite{EnglishTranslationConsolidation2010}
		Given the constraints facing firms, wages (measured in average monthly remuneration) may still be reduced through separations, or by reclassifying workers to occupations with lower nominal wage levels.

@@

		Furthermore, we can rule out the possibility that the distribution of CBAs over time is driven by any particular industry. As Lagos (2021) points out, there is no bunching by certain industries around any particular month for CBAs.
% =================================================================================================
% 													DATA	
% =================================================================================================
	\section{Data}
		\subsection{RAIS}
		The Annual Social Information Report (RAIS) is an administrative dataset containing detailed information about workers and employers at the monthly frequency. The key variables of interest are monthly remuneration, employment status, 6-digit occupation codes (CBO), and 2-digit broad sector classifications from IBGE. We use average monthly remuneration by firm as the main wage variable throughout the study. We use the count of employed workers in a given month as the main employment variable of interest. 
		\begin{figure}
			\centering
			\caption{INSERT RAIS SUM STATS}
			\label{tab:SumStatsRAIS}
		\end{figure}

@@

% =================================================================================================
% 													THEORY	
% =================================================================================================
	\section{Future Research}
	Further analysis is required to pin down what factors are driving the reduction in nominal wages across all firms during the 2015-2016 recession.\footnote{See Figure \ref{fig:NormWages}} One possible explanation is that firms separate from workers during the recession, curtailing the workers' monthly earnings. There is some preliminary evidence of this in Figure \ref{fig:NormEmp}. Another potentially complementary explanation is that firms reclassify workers to occupations with lower implied nominal wages. While we have not yet carried out the analyses to corroborate whether the latter explanation is pertinent, we may have sufficient data to accomplish this in the near future.

	Another salient question is why firms that negotiate wages just after the shock set higher nominal wages than firms negotiating just before Figure \ref{fig:RegressionResults} One possible explanation is that firms in the post-shock group have higher levels of optimism about the future state of the economy and their own future profit margins, which may lead them to set higher wage levels. We may gain access to microdata on firms' own-price and own-cost expectations that would allow us to study the link between individual firms' expectations and negotiated wages as well as hiring and firing decisions.

	Following Olivei and Tenreyo (2007), we may adapt a dynamic stochastic general equilibrium model with uneven staggering of wage contracts and downward rigidity.

	Another approach would be to adapt a DMP model to include inflation expectations. \cite{fukuiTheoryWageRigidity}
